---
title: "Anime"
author: "Richyyl"
date: "26 April 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load packages
```{r}

library(tidyverse)
library(data.table)
library(lubridate)
theme_set(theme_minimal())

```

Read in data from github and tidy for local version
```{r}

tidy_anime <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-04-23/tidy_anime.csv")

tidy_anime_short <- tidy_anime %>%
  select(-synopsis, -background, -title_japanese, -name, -related)

write_csv(tidy_anime_short, "tidy_anime_short.csv")


```

Read in local data
```{r}

tidy_anime_short <- fread("tidy_anime_short.csv") %>%
  mutate(end_year = year(end_date),
         end_decade = end_year %/% 10 * 10,
         start_year = year(start_date),
         start_decade = start_year %/% 10 * 10)

```

There appears to be a slight positive trend between the number of episodes and the rating score, especially with regards to PG-13 and R+ anime.
```{r}

tidy_anime_short %>%
  select(title_english, score, episodes, rating, scored_by) %>%
  filter(episodes > 1,
         scored_by > 1000) %>%
  ggplot(aes(x = episodes, y = score, col = scored_by)) +
  scale_x_log10() +
  geom_point(alpha = 0.5) +
  facet_wrap(~rating) +
  scale_color_gradient2(low = "#e6550d", high = "#31a354", 
                        midpoint = 300000) +
  labs(x = "Number of episodes",
       y = "Rating",
       col = "Ratings")


count(tidy_anime_short, scored_by) %>%
  arrange(desc(scored_by))

```

The median rating for anime has been steadily increasing since the 60s, however there are far more outliers now than in past decades.
```{r}

tidy_anime_short %>%
  select(genre, score, end_decade) %>%
  ggplot(aes(x = end_decade, y = score, group = decade)) +
           geom_boxplot() +
  labs( x = "Ending decade",
        y = "Rating")

```

When was the golden era for anime release?
```{r}

tidy_anime_short %>%
  filter(!is.na(start_decade),
         !is.na(rating)) %>%
  mutate(start_decade = factor(start_decade,
                               ordered = T)) %>%
  group_by(start_decade) %>%
  summarise(n = tally(),
            score = mean(sore, na.rm = T))
  gather(measure, value, c(n, score)) %>%
  ggplot(aes(x = start_decade, y = value, col = measure)) +
  geom_jitter()
  

```


