---
title: "Space race_14-01-2019"
author: "Richyl"
date: "14 January 2019"
output: 
  html_document: 
    df_print: tibble
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = FALSE
)
```

Load packages
```{r}
library(tidyverse)
library(scales)
library(lubridate)
library(countrycode)
theme_set(theme_minimal())
```

Load raw data
```{r}
raw_agencies <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-01-15/agencies.csv") 

agencies_clean <- raw_agencies %>%
  mutate(tstart = str_trim(word(tstart, 1), side = "both"),
         tstart = str_replace(tstart, "\\-|\\*|\\?|s", ""),
         tstop = str_trim(word(tstop, 1), side = "both"),
         tstop = str_replace(tstop, "\\-|\\*", as.character(year(now()))),
         tstop = str_replace(tstop, "\\?", ""),
         agency_type = str_to_title(agency_type),
         Country = countrycode(state_code, "iso2c", "country.name"),
         Country = case_when(state_code == "F" ~ "France",
                             state_code == "I" ~ "Italy",
                             state_code == "J" ~ "Japan",
                             state_code == "SU" ~ "Soviet Union",
                             state_code == "UK" ~ "United Kingodom",
                             state_code %in% c("CYM", "I-ELDO", "I-ESA") ~ "EU",
                             T ~ Country))

raw_launches <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-01-15/launches.csv") 

launches_clean <- raw_launches %>%
  mutate(launch_date = ymd(launch_date),
         make = word(type, 1),
         mission_sub = mission,
         mission = word(mission_sub, 1),
         Country = countrycode(state_code, "iso2c", "country.name"),
         Country = case_when(state_code == "F" ~ "France",
                             state_code == "I" ~ "Italy",
                             state_code == "J" ~ "Japan",
                             state_code == "SU" ~ "Soviet Union",
                             state_code == "UK" ~ "United Kingodom",
                             state_code %in% c("CYM", "I-ELDO", "I-ESA") ~ "EU",
                             T ~ Country))
```

 
Dive into the agencies datset: Longest running, country with most launches, types of agency and type per country
```{r}

oldest_agency <- agencies_clean %>%
  filter(tstart != "") %>%
  mutate_at(.vars = vars(tstart:tstop), .funs = as.numeric) %>%
  mutate(Age = tstop - tstart) %>%
  arrange(desc(Age)) %>%
  ggplot(aes(x = Age)) +
  geom_histogram()

oldest_agency

most_launches <- agencies_clean %>%
  group_by(Country) %>%
  summarise(Launches = sum(count)) %>%
  mutate(Country = fct_reorder(Country, Launches)) %>%
  ggplot(aes(x = Country, y = Launches, fill = Country)) +
  geom_col(show.legend = F) +
  coord_flip() +
  scale_y_continuous(labels = comma_format())

most_launches

ggsave("most_launches.jpg")

agency_launches <- agencies_clean %>%
  group_by(agency_type) %>%
  summarise(Launches = sum(count)) %>%
  mutate(Agency = fct_reorder(agency_type, Launches)) %>%
  ggplot(aes(x = Agency, y = Launches, fill = Agency)) +
  geom_col(show.legend = F) +
  coord_flip() +
  scale_y_continuous(labels = comma_format())

agency_launches



```

Dive into the launches dataset, looking at share of launches by decade, type of rocket and missions
```{r}
launches_by_year <- launches_clean %>%
  count(launch_year) %>%
  ggplot(aes(x = launch_year, y = n, group = 1)) +
  geom_line() +
  geom_point() +
  geom_vline(xintercept = c(1957, 1975), col = "red") +
  labs(title = "Rocket launches rose through the 60s & 70s \n before dropping in the 90s",
       subtitile = "Red lines represent the space race period",
       x = "Launch year",
       y = "Number of launches")

launches_by_year

ggsave("yearly_launches.jpg")

launch_share_by_decade <- launches_clean %>%
  mutate(decade = launch_year %/% 10 * 10,
         agency_type = str_to_title(agency_type)) %>%
  group_by(decade, agency_type) %>%
  summarise(Missions = n()) %>%
  ungroup() %>%
  complete(agency_type, decade, fill = list(Missions = 0)) %>%
  group_by(decade) %>%
  mutate(percentage = round(Missions/ sum(Missions), digits = 2)) %>%
  ggplot(aes(x = decade, y = percentage, fill = agency_type)) +
  geom_area() +
  labs(x = "Decade",
       y = "Percentage of launches",
       fill = "Type of agency") +
  scale_y_continuous(labels = percent_format())

launch_share_by_decade

ggsave("Decade_launches.jpg")

rocket_types <- launches_clean %>%
  count(make) %>%
  top_n(15) %>%
  mutate(make = fct_reorder(make, n)) %>%
  ggplot(aes(x = make, y = n, fill = make)) +
  geom_col(show.legend = F) +
  coord_flip() +
  labs(x = "Rocket make",
       y = "Number of launches made")

rocket_types

mission_types <- launches_clean %>%
  count(mission) %>%
  top_n(15) %>%
  mutate(mission = fct_reorder(mission, n)) %>%
  ggplot(aes(x = mission, y = n, fill = mission)) +
  geom_col(show.legend = F) +
  coord_flip() +
  labs(x = "Mission title",
       y = "Number of attempts")

mission_types
```

where are the non public sector entities located and when did they get on the scene?
```{r}
non_public_agencies <- agencies_clean %>%
  filter(agency_type != "State")

non_public_countries <- non_public_agencies %>%
  count(Country, agency_type) %>%
  ggplot(aes(x = agency_type, y = n, fill = Country)) +
  geom_col(show.legend = T, position = position_dodge2(width = 0.9)) +
  coord_flip() +
  labs(y = "Number of private / startup entities",
       x = "Agency type")

non_public_countries

ggsave("non state agencies.jpg")

non_public_starts <- non_public_agencies %>%
  arrange(tstart) %>%
  mutate(sorter = 1:nrow(.),
         short_name = fct_reorder(short_name, sorter)) %>% 
  filter(tstart != "") %>%
  ggplot(aes(x = tstart, y = short_name, col = agency_type)) +
  geom_point(show.legend = F) +
  labs(x = "Start date",
       y = "Company")

non_public_starts  
```

how successful have startups been in the space race?
```{r}
startup_success <- launches_clean %>%
  filter(agency_type == "startup") %>%
  group_by(launch_year) %>%
  count(category) %>%
  complete(launch_year, fill = list(n = 0)) %>%
  ungroup() %>%
  mutate(launch_year = factor(launch_year)) %>%
  ggplot(aes(x = launch_year, y = n, col = category, group = category)) +
  geom_jitter() +
  geom_smooth(se = F, method = "auto", span = 1) +
  labs(y = "Number of launches",
       x = "Launch year",
       col = "Mission outcome",
       title = "Startups have shown increasing launch success,
       since initial launches")

ggsave("startup_success.jpg")

```

